---
title: "Recommendations"
author: "Vladimir Mikheev"
date: "06 07 2021"
output:
  html_document:
    code_download: yes
    fontsize: 8pt
    highlight: textmate
    number_sections: no
    theme: flatly
    toc: yes
    toc_float:
      collapsed: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(dplyr)
library(data.table)
library(readr)
library(lme4)
library(broom)
library(ggplot2)
library(tidyr)
library(arm)
library(ggpmisc)


remove_outliers <- function(x, na.rm = TRUE, ...) {
  qnt <- quantile(x, probs=c(.25, .75), na.rm = na.rm, ...)
  H <- 1.5 * IQR(x, na.rm = na.rm)
  y <- x
  y[x < (qnt[1] - H)] <- NA
  y[x > (qnt[2] + H)] <- NA
  y
}

```


```{r, include=FALSE}
wf <- 'C:/Users/Vladimir/YandexDisk/Достижения/Criteo/ecommerce_dataset.csv'
#df <- fread('unzip -cq C:/Users/Vladimir/YandexDisk/Достижения/Criteo/ecommerce_dataset.zip') 
df <- fread(wf) %>% mutate(
  Datetime =as.POSIXct(as.numeric(as.character (Timestamp)), origin="1970-01-01", tz="GMT"), 
                           `Product Price` = as.numeric(`Product Price`), 
                           `Product quantity` = as.numeric(`Product quantity`),
  Date = as.Date(Date, "%d/%m/%y"), 
  Weekdays = weekdays(Date),
  `Device type` = if_else(`Device type` == 'Unknown', "Android app", `Device type`),
  `Browser family` = if_else(`Browser family` == 'other', "Android app", `Browser family`)
  ) %>% 
  mutate(`Device group` = case_when(
    `Device type` %in% c('Android - Smartphone', 'iPhone', 'Mobile - Other') ~ 'Smartphone',
    `Device type` %in% c('Android - Tablet', 'iPad') ~ 'Tablet',
    `Device type` == 'Android app' ~ 'Android app',
    `Device type` == 'Desktop' ~ 'Desktop'
  ))
  
df$`Event type` = factor(df$`Event type`, levels=c('Listing', 'Product', 'Basket', 'Sales'))
df$`Existing client` = as.factor(df$`Existing client`)
df$`Weekdays` = factor(df$`Weekdays`, levels = c('понедельник','вторник','среда','четверг','пятница','суббота','воскресенье')) 


```


```{r, include=FALSE}
table(df$`Browser family`)
table(df$`User location`)
table(df$`Event type`)
table(df$`Device type`)
table(df$`Environment`)
```



# 1. Timing
## 1.1. Hour

Let's observe at what time sales occur

```{r, echo=FALSE}
df %>% filter(`Event type` %in% c('Sales', 'Basket')) %>% 
  group_by(Hours, `Event type`) %>% 
  mutate(Revenue = sum(`Product Price` * `Product quantity`), na.rm = TRUE) %>% 
  ggplot(aes(x = Hours, y = Revenue )) + geom_line() + 
  facet_grid(~`Event type`) +
  stat_peaks(colour = "red",  ignore_threshold = 0.99)  +
  stat_peaks(color = "red", geom = "rug", sides = "b", ignore_threshold = 0.99)+
  ggtitle('Distribution of basket and sale events over a day') + theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))

```

The purchasing spans from 9 o'clock to 23 o'clock. The most popular time for sales is 20 o'clock.

```{r, echo=FALSE}

df %>% 
  group_by(Hours, `Event type`) %>% 
  mutate(Visits = n()) %>% 
  ggplot(aes(x = Hours, y = Visits )) + geom_line() + 
  facet_grid(~`Event type`) +
  stat_peaks(colour = "red",  ignore_threshold = 0.99)  +
  stat_peaks(color = "red", geom = "rug", sides = "b", ignore_threshold = 0.99) +
  ggtitle('Distribution of events over a day') + theme_minimal()  +
  theme(plot.title = element_text(hjust = 0.5))

```

 
Product visiting, basketing and purchasing happens mostly at 20 o'clock, while listing is uniformly distrusted across day with maximum at 16 o'clock. 


**Recommendation:** the perfect timing for advertisment is between 16 to 20 o'clock  

## 1.2. Day of week

```{r, echo=FALSE}
df %>% filter(`Event type` == 'Sales') %>% 
  group_by(Weekdays) %>% 
  summarise(Visits = n()) %>% 
  ggplot(aes(x = Weekdays, y = Visits, fill =  Weekdays, label = Visits)) + 
  geom_col() + geom_text( position = position_dodge(width = 0.9), vjust = -0.5) +
  ggtitle('Distribution of sale events over an average week') + 
  labs(y = "Number of sales", x = 'Weekdays', fill = 'Number of sales')  + theme_minimal() + 
  theme(plot.title = element_text(hjust = 0.5))  

```
```{r}
# proof

#1. упорядочить по дням
#2. ти тест или анова чтобы доказать значимость различий между днями.
#m <- lm()
```

**Recommendation:** the most sailing days are Friday, Saturday and Sunday, so advertising campaign could be more effective in this days.


# 2. Browser

```{r, echo=FALSE}

df  %>% filter(`Event type` == 'Sales', `Browser family` != 'NULL', `Browser family` != 'other') %>%
  group_by(`Browser family`) %>% 
  summarise(Usage = n()) %>% arrange(desc(Usage)) %>% 
  mutate(`Browser family` = case_when(
    Usage < 100  ~ 'small',
    Usage > 100 ~ `Browser family`)) %>% 
  filter(`Browser family` != 'small') %>% 
  ggplot(aes(y = Usage, x = reorder(`Browser family`, Usage), label = Usage , fill = reorder(`Browser family`, desc(Usage)))) + 
  geom_col() +
  geom_text( position = position_dodge(width = 0.9), vjust = -0.5) +
  ggtitle('Distribution of browser families') + 
  labs(y = "Number of sales", x = 'Browser family', fill = 'Browser family')  + theme_minimal()  + 
  theme(plot.title = element_text(hjust = 0.5))  
```

Let's also observe browser usage through different device groups.

```{r, echo=FALSE}
 df  %>% filter(`Event type` == 'Sales', `Browser family` != 'NULL', `Browser family` != 'other') %>%
  group_by(`Browser family`, `Device group`) %>% 
  summarise(Usage = n()) %>% arrange(desc(Usage)) %>% 
  mutate(`Browser family` = case_when(
    Usage < 90  ~ 'small',
    Usage > 90 ~ `Browser family`)) %>% 
  filter(`Browser family` != 'small')  %>% 
  ggplot(aes(x = Usage, y = reorder(`Browser family`, Usage), label = Usage , fill = reorder(`Browser family`, desc(Usage)))) + 
  geom_col() +
  geom_text( position = position_dodge(width = 0.9)) + 
  facet_wrap(~`Device group`) +
  ggtitle('Distribution of browser families') + 
  labs(x = "Number of sales", y = 'Browser family', fill = 'Browser family')   + theme_minimal()   + 
  theme(plot.title = element_text(hjust = 0.5))


```
**Recommendation:**

Different browsers render content in their own way. So during advertisement campaign company should adjust to the following browsers as soon as they affect the majority of sales:

 - Chrome, Safari and Firefox for desktop; 
 - Mobile Safari and Chrome Mobile for smartphones;
 - Mobile Safari and Chrome for Tablets. 


```{r, include=FALSE}
df  %>% filter(`Event type` == 'Sales', `Browser family` != 'NULL', `Browser family` != 'other') %>%
  group_by(`Browser family`) %>% 
  summarise(Usage = n()) %>% arrange(desc(Usage)) %>% 
  mutate(`Browser family` = case_when(
    Usage < 100  ~ 'small',
    Usage > 100 ~ `Browser family`)) %>% 
  filter(`Browser family` != 'small') %>% distinct( `Browser family`) -> list_bro
list_bro <- as.list(list_bro$`Browser family`)
```

```{r, echo=FALSE}
# position dodge?

df  %>% filter( `Browser family` != 'NULL', `Browser family` != 'other') %>%
  group_by(`Browser family`, `Event type`) %>% 
  summarise(Usage = n()) %>% 
  filter(`Browser family` %in%  list_bro) %>% 
  ggplot(aes(x = Usage, y = reorder(`Browser family`, Usage), label = Usage , fill = reorder(`Browser family`, Usage))) + 
  geom_col() +
  geom_text( position = position_dodge(width = 0.9)) +
  ggtitle('Distribution of browser families over actions') + 
  labs(x = "Number of sales", fill = 'Browser family')  + facet_wrap(~`Event type`)+ theme_minimal()  +
  theme(plot.title = element_text(hjust = 0.5))   
```


# 2. User location

```{r, echo=FALSE}
df %>% filter(`User location` != "Unknown") %>%  filter(`Event type` == 'Sales') %>% 
  group_by(`User location`, Date) %>% 
  mutate(Usage = n()) %>% 
  ggplot(aes(x = Date, y = Usage, color = `User location`)) + geom_col(show.legend = FALSE) + facet_wrap(~`User location`)  +
  ggtitle('Distribution of browser families over actions') + theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))   
```

```{r, echo=FALSE}
df %>% filter(`User location` != "Unknown", `Event type` == 'Sales') %>%  
  group_by(`User location`) %>% 
  summarise(Usage = n()) %>% 
  ggplot(aes(x = `User location`, y = Usage, fill = `User location`, label = Usage)) +
  geom_col() +
  geom_text(position = position_dodge(width = 0.9), vjust = -0.5)   + 
  theme_minimal() + ggtitle('Annual number of sales actions from inland and outland') + 
  theme(plot.title = element_text(hjust = 0.5))

```

```{r, echo=FALSE}
df %>% filter(`User location` != "Unknown", `Event type` == 'Sales') %>%  
  group_by(`User location`, `Product category`) %>% 
  summarise(Usage = n()) %>% 
  ggplot(aes(x = `User location`, y = Usage, fill = `User location`, label = Usage)) +
  geom_col(show.legend = FALSE) +
  geom_text(position = position_dodge(width = 0.9), vjust = -0.5)   + 
  facet_grid(~`Product category`) +
  theme_minimal() + ggtitle('Annual number of sales actions from inland and outland') + 
  theme(plot.title = element_text(hjust = 0.5))
```



**Recommendation:** only 10 % of sales occur from outside, so outside advertisment should not be a main focus of a campaign.


# 3. Existing client

```{r, echo=FALSE}

df %>% filter(`Event type` == 'Sales') %>%  
  group_by(`Existing client`) %>% 
  summarise(Usage = n()) %>% 
  ggplot(aes(x = `Existing client`, y = Usage, fill = `Existing client`, label = Usage)) +
  geom_col(show.legend = FALSE) +
  geom_text(position = position_dodge(width = 0.9), vjust = -0.5)   + 
  theme_minimal() + ggtitle('Annual number of sales actions from new and exising clients') + 
  theme(plot.title = element_text(hjust = 0.5)) + scale_fill_hue(name = "Type of client", labels=c("New", "Existing")) + scale_x_discrete(labels=c("New", "Existing"))
```

```{r, echo=FALSE}

df %>%  filter(`Event type` == 'Sales') %>% group_by(`Cross-device user ID`) %>% mutate(Revenue = sum(`Product Price` * `Product quantity`)) %>% group_by(`Existing client`) %>% summarise(`Total Revenue` = sum(Revenue), `Mean Revenue` = mean(Revenue),`Median Revenue` = median(Revenue), `Mean Product Quantity` = mean(`Product quantity`)) 
```
```{r, echo=FALSE}
df %>%  filter(`Event type` == 'Sales') %>% group_by(`Existing client`, Date) %>% 
  mutate(Revenue = sum(`Product Price` * `Product quantity`)) %>% 
  ggplot(aes(x = Date, y = Revenue, fill = `Existing client`)) + geom_col(show.legend = FALSE) + 
  facet_wrap(~`Existing client`, labeller = as_labeller(c(`0` = "New", `1` = "Existing"))) +
  ggtitle('Distribution of sales over type of purchasers') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))   
```

New clients buy significantly more 

# 3. Product category
, echo=FALSE
```{r}
df %>%  filter(`Event type` == 'Sales') %>% group_by(`Product category`) %>% 
  summarise(Revenue = sum(`Product Price` * `Product quantity`)) %>% 
  ggplot(aes(x =`Product category`, y = Revenue, label = Revenue, fill= `Product category`))+   
  geom_col() +  
  geom_text(position = position_dodge(width = 0.9), vjust = -0.5)   + 
  theme_minimal() + ggtitle('Annual Revenue for product categories') + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r, echo=FALSE}
df %>%  filter(`Event type` == 'Sales') %>% group_by(`Product category`) %>% 
  summarise(Quantity = sum(`Product quantity`)) %>% 
  ggplot(aes(x =`Product category`, y = Quantity, fill =`Product category`, label = Quantity))+  
  geom_col() +  
  geom_text(position = position_dodge(width = 0.9), vjust = -0.5)   + 
  theme_minimal() + ggtitle('Annual quantity of saled products') + 
  theme(plot.title = element_text(hjust = 0.5))
```


```{r, echo=FALSE}
df %>%  filter(`Event type` == 'Sales') %>% group_by(`Product category`, Date) %>% 
  mutate(Quantity = sum(`Product quantity`)) %>% 
  ggplot(aes(x = Date, y = Quantity, color = `Product category`))+ 
  geom_col(show.legend = FALSE) + facet_wrap(~`Product category`) + 
  ggtitle('Annual sales events by  product category') + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r, echo=FALSE}
df %>%  filter(`Event type` == 'Sales') %>% group_by(`Product category`, Date) %>% 
  mutate(Revenue = sum(`Product Price` * `Product quantity`)) %>% 
  ggplot(aes(x = Date, y = Revenue, color = `Product category`))+ 
  geom_col(show.legend = FALSE) + facet_wrap(~`Product category`)+ 
  ggtitle('Annual revenue by product category') + 
  theme(plot.title = element_text(hjust = 0.5))
```

**Recommendations:**

  1. packages and small item were saled in the same quantity, but revenue from packages is highest among all products and by two times higher than from small items 
  2. Advertising of two leading products, packages and small items, should give the highest revenue.
  3. In November-December promotion of  packages and medium items is preferable.
  4. In April promotion of  packages and small items is preferable.
  
  
# 5. Sales funnel

```{r, echo=FALSE}
df %>% 
  group_by(`Event type`) %>% 
  summarise(Visits = n()) 
```

```{r, echo=FALSE}
tab <- df %>% 
  group_by(`Cross-device user ID`) %>%   
  summarise(`Number of actions` = n_distinct(`Event type`), `Listed?` = if_else('Listing' %in% `Event type`, 1, 0), `Sales?` = if_else('Sales' %in% `Event type`, 1, 0)) %>% filter(`Listed?` != 0) %>% group_by(`Sales?`) %>%  summarise(`Number of those who listed` = n())
tab

(tab$`Number of those who listed`[2] *100) /  tab$`Number of those who listed`[1]


tab <- df %>% 
  group_by(`Cross-device user ID`) %>%   
  summarise(`Number of actions` = n_distinct(`Event type`), `Product?` = if_else('Product' %in% `Event type`, 1, 0), `Sales?` = if_else('Sales' %in% `Event type`, 1, 0)) %>% filter(`Product?` != 0) %>% group_by(`Sales?`) %>%  summarise(`Number of those who visited product page` = n())
tab

(tab$`Number of those who visited product page`[2] *100) /  tab$`Number of those who visited product page`[1]
```

That means that

  1. Among whose who listed only 17% purchased a product, while among those who visited product page - 19%. Majority of shoppers do not list and navigate directly to the required product. 
  2. Кажется нужен пивалью











